---
title: "Measuring Ecological Intelligence on Business Students"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: tango
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = "./../docs/index.html") })
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

library(openxlsx)
library(dplyr)
library(kableExtra)
library(psych)
library(ggplot2)
library(Hmisc)
library(ggcorrplot)
library(lavaan)
library(GGally)
library(ggplot2)
library(gridExtra)
library(qqplotr)
```

## Intelligence Scale

Ecological intelligence aims to develop social and environmental responsability and awareness, to think critically, to pursue cooperative learning, and to bring about behavioral change in the long-term.

Ecological intelligence should have a holistic perspective because there are seen and unseen webs among biotic and abiotic factors in the world. According to the model of Okur-Berberoglu (2020), social intelligence and economy are subsets of ecological intelligence. Social intelligence refers to social responsibilities of people in terms of sustainability. Economy should be based on sustainable development instead of explotation of environmental and human resources.

Okur-Berberoglu (2020) developed a scale for ecological intelligence. The questionnaire consists of 12 questions based on a 5 Likert scale that were scored as 1-Completely disagree, 2-Partly disagree, 3-Not sure, 4-Partly agree, and 5-Completely agree. The questions were the following.\
1. I get my full water bottle while leaving home.\
2. I wonder about increasing of cancer events in industrial areas.\
3. Degenerated environmental conditions can cause negative effect on mental health.\
4. What a pitty!, people think that technology can solve every problem, althought technology cannot produce one gram organic honey.\
5. The reflection of environmental problems can be seen at the same time, in succession, and more than one area.\
6. I try to pattern people who have positive ecologic behavior in my social life.\
7. I prefer to buy local vegetables and fruits.\
8. I believe that one of the ways of fighting with obesity is environmental education.\
9. I have remorse to know some goods I buy, are produced by explotation of human work.\
10. One of the reasons of immigration of rural people, is the job lost due to global firms.\
11. Global firms prevent local producer to have profit.\
12. People, to immigrate to another place, are cheaper laborer sources for big companies.

This scale was based on Turkey and Turkish lifestyle, as the author comment in his work this scale should the tested and understood by other researchers.

## Survey Data

```{r Read data, echo=TRUE}
data_file_path <- here::here("data", "Ecologial Scale - Raw Data.xlsx")
#dfenc <- read.xlsx(xlsxFile = "./data/Ecologial Scale - Raw Data.xlsx", 
#                     sheet = "Ecological Inteligence Scale", colNames=TRUE) %>%
dfenc <- read.xlsx(xlsxFile = data_file_path) %>%
  mutate(Username = seq(1,n())) %>%
  na.omit()
colnames(dfenc) <- c("date", "student", "campus", "genre", "nationality", "age", paste0("holis", 1:5), paste0("social", 1:4), paste0("econ",1:3), "change", "argument") 
dfenc$campus <- substr(dfenc$campus, 8, 50)
for (i in 7:18) {
  dfenc[,i] <- recode(dfenc[,i], `Completely agree`=5,
                      `Partly agree`=4,
                      `Not sure`=3,
                      `Partly disagree`=2,
                      `Completely disagree`=1)
}
```

```{r, echo=TRUE}
dfdemo <- select(dfenc, student:age)

countvals <- function(x, categ) {return(sum(x==categ))}
agetags <- sort(unique(dfenc$age))

tbdesc <- dfdemo %>%
  group_by(campus) %>%
  summarise(N = n(),
            Hombre = countvals(genre, "Hombre"),
            Mujer = countvals(genre, "Mujer"),
            USA = countvals(nationality, "Estadounidense"),
            Mexico = countvals(nationality, "Mexicana"),
            Otro = countvals(nationality, "Otra"),
            E21_22 = countvals(age, agetags[1]),
            E23_24 = countvals(age, agetags[2]),
            E25omas = countvals(age, agetags[3])) 

kbl(tbdesc) %>%
  kable_classic(full_width=FALSE) %>%
  add_header_above(c(" "=2, "Sex"=2, "Country"=3, "Age"=3))
```

## Descriptive

Now we compute the indicators and its statistics.

```{r, echo=TRUE}
dfenc$campus[dfenc$campus == "Aguascalientes"] <- "Otro"
dfenc$campus[dfenc$campus == "Morelia"] <- "Otro"

options(width=200)
dfenc <- dfenc %>%
  mutate(holis = (holis1+holis2+holis3+holis4+holis5)/5,
         social = (social1+social2+social3+social4)/4,
         econ = (econ1+econ2+econ3)/3)
psych::describe(select(dfenc,holis:econ, change))
```

Let us obtain the correlation matrix and their corresponding p-values.

```{r, echo=TRUE}
rcorr(as.matrix(select(dfenc,holis1:econ3)))
```

## Cronbach's alpha

When items are used to form a scale they eed to have internal consistency. The items should all measure the same thing, so they should be correlated with on another. A useful coefficient for assessing internal consistency is Cronbach's alpha. The formula is: $$
\alpha = \frac{k}{k-1} \left( 1-\frac{\sum s_i^2}{s_T^2} \right)
$$ where $k$ is the number of items, $s_i^2$ is the variance of the ith item and $s_T^2$ is the variance of the total score formed by summing all the items. If the items are all independent, then $\alpha =0$. If the items are all identical and so perfectly correlated, then $\alpha = 1$.

```{r, echo=TRUE}
psych::alpha(select(dfenc, holis1:econ3))
psych::alpha(select(dfenc, holis1:holis5))
psych::alpha(select(dfenc, social1:social4))
psych::alpha(select(dfenc, econ1:econ3))
```

We also visualize the correlations with a heat map. We note that the correlations are low, so the instrument would not capture the construct properly.

```{r, echo=TRUE}
p.mat <- cor_pmat(select(dfenc,holis1:econ3))
ggcorrplot(cor(select(dfenc,holis1:econ3)), 
           type="lower", 
           lab=TRUE,
           lab_size=3,
           insig="blank",
           p.mat=p.mat,
           sig.level=0.10)
```

Let us compare the dimensions by campus.

```{r, echo=TRUE}
stkfacts <- data.frame()
namefacts <- c("Holistic", "Social", "Econ")
tagfacts <- c("holis", "social", "econ")
for (i in 1:3) {
  dftmp <- data.frame(student = dfenc$student,
                      campus = dfenc$campus,
                      fact=rep(namefacts[i],nrow(dfenc)),
                      value = dfenc[, tagfacts[i]])
  stkfacts <- rbind(stkfacts, dftmp)
}

ggplot(stkfacts, aes(y=fact, x=value, fill=campus)) +
  geom_boxplot() +
  theme(axis.title.y = element_blank())
```

## Kaiser-Meyer-Olkin (KMO) test

The KMO test is a statistical measure to determine how suited data is for factor analysis. The test measures sampling adequacy for each variable in the model and the complete model. The statistic is a measure of the proportion of variance among variables that might be common variance.

The MSA (Measure of Sampling Adequacy) criterion is calculated and return values between 0 and 1. $$
\hbox{MSA} =  \frac{\sum_{j\neq k} r_{jk}^2}{\sum_{j\neq k} r_{jk}^2 + \sum_{j\neq k} p_{jk}^2}
$$ Here $r_{jk}$ is the correlation between the variable in question and another, and $p_{jk}$ is the partial correlation. KMO values between 0.8 and 1 indicate the sampling is adequate. KMO values less than 0.6 indicate the sampling is not adequate and that remedial action should be taken. In contrast, others set this cutoff value at 0.5. A KMO value close to zero means that there are large partial correlations compared to the sum of correlations. In other words, there are widespread correlations which would be a large problem for factor analysis.

```{r, echo=TRUE}
cor_matrix <- cor(select(dfenc,holis1:econ3))
KMO(cor_matrix)
```

## Bartlett's test of sphericity

The Bartlett's test is used to determine whether or not the data come from a multivariate normal distribution.

Let us suppose that $x_1, x_2, \dots$ are random vectors en $R^p$ having a multivariate normal distribution with unknown mean vector $\mu$ and unknown covariance matrix $\Sigma$. The hypothesis are\
$H_o: \Sigma = \sigma I_p$\
$H_a: \Sigma \neq \sigma I_p$, $\Sigma$ positive definite\
$\alpha = 0.05$\
The test statistic is $$
T_0 = \frac{(\hbox{tr}\ S)^p}{|S|}.
$$

```{r, echo=FALSE}
cor_matrix <- cor(select(dfenc,holis1:econ3))
cortest.bartlett(cor_matrix, n = nrow(dfenc))
```

## Confirmatory analysis

```{r, echo=TRUE}
dfitems <- select(dfenc, holis1:econ3)
mla <- 'f1 =~ holis1 + holis2 + holis3 + holis4 + holis5
        f2 =~ social1 + social2 + social3 + social4
        f3 =~ econ1 + econ2 + econ3'
threef <- cfa(mla, data=dfitems, std.lv=TRUE)
summary(threef, fit.measures=TRUE, standardized=TRUE)
```

## Multiple linear regression

```{r, echo=TRUE}
p.mat <- cor_pmat(select(dfenc,change, holis:econ))
ggcorrplot(cor(select(dfenc,change, holis:econ)), 
           type="lower", 
           lab=TRUE,
           lab_size=3,
           insig="blank",
           p.mat=p.mat,
           digits=4,
           sig.level=0.10)

ggpairs(select(dfenc,holis:econ, change), aes(alpha=0.1))
```

```{r, echo=TRUE}
model <- lm(change ~ campus + holis + social + econ, data=dfenc)
summary(model)

# Residual analysis
p1 <- ggplot(mapping = aes(sample = residuals(model))) + 
  stat_qq_point(size = 1, color="black") + 
  stat_qq_line(color="red") + 
  labs(x="Change", y = "Residuals")
p2 <- ggplot(model, aes(x=fitted(model), y=residuals(model))) +
  geom_point() +
  #geom_point(group=model$model$change, color=model$model$change) + 
  labs(x="Fitted values", y="Residuals") +
  theme(legend.position = "top")
grid.arrange(p1, p2, nrow=1)

# Kolmogorov-Smirnov test
ksout <- ks.test(residuals(model), "pnorm", mean=0, sd=sd(residuals(model))) 
print(ksout)
```

## References

Okur-Berberoglu, E. (2020). An ecological intelligence scale intended for adults. World Futures, 76(3), 133-152.

Yves Rosseel (2012). lavaan: An R Package for Structural Equation Modeling. Journal of Statistical Software, 48(2), 1-36. URL <http://www.jstatsoft.org/v48/i02/>
